Malacology: A Programmable Storage System Built on Ceph

### 2016-05-10 paper/paper.tex
Edits to section 4 intro and 4.1

To illustrate the benefits and challenges of this approach we examine
the programmability of Ceph~\cite{weil_ceph_2006}, the increasingly
popular, production-level open-source distributed storage system.
Something of a storage swiss army knife, Ceph supports file, block, and
object interfaces simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous 
Distributed Object Storage (RADOS) system is a cluster of storage
devices (OSDs) that provide Ceph with data durability and integrity
using replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
% By introducing
% programmability concepts into Ceph, we can build new services by
% carefully exposing internal storage services to applications.

@@

So called ``bolt-on'' services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on'' services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{corfu}. Without going into greater detail, the CORFU protocol, by separating the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
``sequencer'' service. This high-performance in-memory counter is used to assign a log
position to each client that appends to the log. The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means). Thus, it is the volatility of the sequencer that enables high performance. A ``bolt-on'' option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the preconceived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

@@

\section{Re-Usable Components in Ceph}\label{re-usable-components-in-ceph}

@@

Ceph is a production-quality distributed storage system and in this section we
touch on some of the subsystems it uses to provide a general-purpose
storage system. In the next section, we describe how Malacology exposes these
subsystems in order to enable the creation of new services.

@@

% straw man example
Keeping track of the state of a distributed system is an essential part of any 
successful service and a necessary component in order to diagnose and detect 
failures, when they occur. In the case of Ceph, a consistent view of cluster 
state among server daemons and clients is critical
to provide strong consistency guarantees to clients. Ceph maintains cluster
state information in per-subsystem ``maps'' (e.g. OSD map, MDS map) that
record membership and status information. A PAXOS-based monitoring service is
responsible for integrating state changes into cluster maps, and responding to
requests from out-of-date clients. As a fundamental building block of many
system designs, consensus abstractions such as PAXOS are a common technique
for maintaining consistent data versions, and are a useful system to expose.
While Ceph does not expose this service directly, it does allow limited use by
clients through a key-value service intended for low-volume, small size
configuration information.

@@

Although this feature is powerful, the current implementation has drawbacks.
Supporting only C/C++ for object interface developers, Ceph requires
distribution of compiled binaries for the correct architecture, adding a
large barrier of entry for developers and system administrators. Second,
having no way to dynamically unload modules, any changes require a full
restart of an OSD which may have serious performance impacts due to a loss
of cached data. And finally, the security limitations of the framework
limit the use of object classes to all but those with administrative level
access and technique expertise.

@@

\subsection{Load Balancing}

@@

In this section we describe how Malacology exposes and builds upon existing
services within Ceph. While we strive for implementation independent
approaches within Malacology, our prototype implementation makes heavy use of
the embedded language Lua to enable dynamic features without relying on
complex and heavy-weight code generation techniques.

{\bf Lua integration.}
Supporting dynamic reconfiguration and programmabiilty is critical in real
systems that must adapt to a changing environment where potentially
conflicting policies arise. We've adopted Lua as a scripting language to
implement dynamic components in Malacology.  Lua is a portable, embedded
scripting language that offers superious performance and productivity
trade-offs, including a JIT-based implementation that is well known for near
native performance. Lua has been used extensively in game engines, and systems
research \cite{neto:dls14-luaos}, including storage systems where it has been
effectively used both on
\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo} and off
\cite{sevilla:sc15-mantle} the performance critical path.

@@

{\bf Balancing Policy.}
Avoiding hard-coded balancing policies are important. First, it is difficult,
nay impossible, to select a one-size-fits-all policy. System administrators
may want to partition load based on SLAs, etc... Applications are often in
a better position to know what to do.

@@

Lua integration. Overhead?

Although these OSD interfaces are powerful, the current implementation
has drawbacks. OSD interfaces are written in C/C++ and compiled into a
shared library, so the developer must account for different target
architectures. Second, C/C++ provides more functionality than is
necessary since the snippets of interface code mainly set policies and
perform simple operations. Finally, developing these interfaces in C/C++
has high overhead. The developer must learn how the OSD dynamically

>>>>

To illustrate the benefits and challenges of this approach we examine
the programmability of Ceph~\cite{weil:osdi2006-ceph}, the increasingly
popular, production-level open-source distributed storage system.
Something of a storage swiss army knife, Ceph supports file, block, and
object interfaces simultaneously in a single cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous 
Distributed Object Storage (RADOS) system is a cluster of storage
devices (OSDs) that provide Ceph with data durability and integrity
using replication, erasure-coding, and scrubbing~\cite{weil_rados_2007}. 
% By introducing
% programmability concepts into Ceph, we can build new services by
% carefully exposing internal storage services to applications.

@@

So called ``bolt-on'' services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on'' services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{balakrishnan_corfu_2012}. Without going into greater detail, the CORFU protocol, by separating the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
``sequencer'' service. This high-performance in-memory counter is used to assign a log
position to each client that appends to the log. The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means). Thus, it is the volatility of the sequencer that enables high performance. A ``bolt-on'' option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the preconceived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

@@

\section{Components of Ceph Suitable for Reusability}\label{re-usable-components-in-ceph}

@@

Ceph is a production-quality distributed storage system and in this section we
touch on some of the subsystems it uses to provide a general-purpose
storage system. In the next section, we describe how Malacology exposes these
subsystems in order to make them reusable and enable the creation of new services on top of them.

@@

% straw man example
Keeping track of the state of a distributed system is an essential part of any 
successful service and a necessary component in order to diagnose and detect 
failures, when they occur. In the case of Ceph, a consistent view of cluster 
state among server daemons and clients is critical
to provide strong consistency guarantees to clients. Ceph maintains cluster
state information in per-subsystem ``maps'' (e.g. OSD map, MDS map) that
record membership and status information. A PAXOS-based monitoring service is
responsible for integrating state changes into cluster maps, and responding to
requests from out-of-date clients. As a fundamental building block of many
system designs, consensus abstractions such as PAXOS are a common technique
for maintaining consistent data versions, and are a useful system to expose.

While Ceph does not expose this service directly, it does allow limited use by
clients through a key-value service intended for low-volume, small size
configuration information. Unfortunately, while a key-value service is
generally useful, it doesn't provide many of the useful services hidden within
the monitoring framework that would be required for applications with more
demanding requirements.

@@

@@

\subsection{Metadata Load Balancing}

@@

In this section we describe how Malacology builds upon existing
services within Ceph. The main idea behind Malacology is to take the subsystems 
described in the previous section and expose APIs on top of them that allow 
users to dynamically inject code, allowing them to customize the behavior of 
these components. Supporting dynamic reconfiguration and programmability 
efficiently is critical in real systems that must adapt to a changing 
environment where potentially
conflicting policies arise. We've adopted Lua as the dynamic runtime  to
implement dynamic components in Malacology.

There are multiple reasons that make Lua an attractive runtime for implementing 
Malacology. Lua is a portable, embedded
scripting language that offers superious performance and productivity
trade-offs, including a JIT-based implementation that is well known for near
native performance. Additionally, Lua has been used extensively in game engines, 
and systems research \cite{neto:dls14-luaos}, including storage systems where it 
has been effectively used both on
\cite{grawinkel:pdsw2012-lua,watkins2013:bdmc13-in-vivo,geambasu_comet_2010} and 
off
\cite{sevilla:sc15-mantle} the performance critical path. Finally, the 
flexibility of the runtime allows to sandbox applications in order to address 
security and performance concerns. We now describe how Malacology exposes each 
subcomponent in Ceph.

The implementation of Ceph's object abstraction, although powerful, has some 
drawbacks.
Supporting only C/C++ for object interface developers, Ceph requires
distribution of compiled binaries for the correct architecture, adding a
large barrier of entry for developers and system administrators. Second,
having no way to dynamically unload modules, any changes require a full
restart of an OSD which may have serious performance impacts due to a loss
of cached data. And finally, the security limitations of the framework
limit the use of object classes to all but those with administrative level
access and technique expertise.

By embedding the Lua runtime, Malacology makes the base RADOS API available in a 
dynamic environment, resulting in a object storage API with economy of 
expression, which at the same time provides the full set of features of the base 
object class. New object interfaces that are expressed in thousands of lines of 
code can be implemented in approximately an order of magnitude less 
code~\cite{geambasu_comet_2010}.

@@

{\bf Balancing Policy.}

Avoiding hard-coded balancing policies is important. First, it is difficult,
nay impossible, to select a one-size-fits-all policy. System administrators
may want to partition load based on SLAs, etc... Applications are often in
a better position to know what to do.

@@
